\documentclass{article}
\usepackage{algorithm,algorithmic}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage[margin=1.00in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tikz}

% new theorems
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}[theorem]{Example}

\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}

% new commands and math operators
\newcommand*\conj[1]{\overline{#1}}
\newcommand{\iu}{{i\mkern1mu}}
\newcommand\abs[1]{\left|#1\right|}
\newcommand\norm[1]{\left\Vert#1\right\Vert}
\newcommand\re[1]{\operatorname{Re}\left(#1\right)}
\newcommand\tr[1]{\operatorname{tr}\left(#1\right)}

\title{A Note to Panos:\\
\emph{\large{Q Numerical Range of the Graph Laplacian}}}
\author{Thomas R. Cameron}
\date{\today}

\begin{document}
\maketitle
\abstract{
This note introduces the Q numerical range of the graph Laplacian and its use for characterizing certain digraphs. 
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                    				Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}	
Let $\Gamma$ be a directed graph (digraph) with non-negative weights and let $L$ be the graph Laplacian of $\Gamma$. 
Furthermore, let $e$ denote the all ones vector. 
Then, the algebraic connectivity of $\Gamma$ is defined as follows~\cite{Wu2005-1}:
\[
\alpha(\Gamma)=\min_{x\in S}x^{T}Lx,
\]
where
\[
S=\left\{x\in\mathbb{R}^{n}\colon x\perp e,\norm{x}=1\right\}.
\]
Another related and useful quantity is the following:
\[
\beta(\Gamma)=\max_{x\in S}x^{T}Lx.
\]
It is important to note that both $\alpha(\Gamma)$ and $\beta(\Gamma)$ are invariant under re-ordering of vertices of $\Gamma$ since the space $S$ is invariant under permutation.

Let $Q$ be an orthonormal matrix whose columns span $S$.
Then, we have
\[
\alpha(\Gamma)=\min_{\norm{Qx}=1}x^{T}Q^{T}LQx
\]
and
\[
\beta(\Gamma)=\max_{\norm{Qx}=1}x^{T}Q^{T}LQx.
\]
Furthermore, let $W(A)$ denote the numerical range (field of values) of a complex matrix $A$, as defined in~\cite{Horn1991}.
Then, the Hermitian part of $A$, denoted $H(A)=\frac{1}{2}(A+A^{*})$, satisfies the following:
\[
W(H(A))=\re{W(A)}.
\]
Moreover, the end points of $\re{W(A)}$ are well-known to be the minimum and maximum eigenvalues of $H(A)$. 
Therefore, we have
\begin{equation}\label{eq:alpha}
\alpha(\Gamma)=\lambda_{\text{min}}\left(\frac{1}{2}Q^{T}(L+L^{T})Q\right)
\end{equation}
and
\begin{equation}\label{eq:beta}
\beta(\Gamma)=\lambda_{\text{max}}\left(\frac{1}{2}Q^{T}(L+L^{T})Q\right).
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                    				The Q Numerical Range of the Graph Laplacian
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Q Numerical Range of the Graph Laplacian}
Motivated by~\eqref{eq:alpha} and~\eqref{eq:beta}, we are interested in studying the numerical range of the matrix $Q^{T}LQ$, where $L$ is the graph Laplacian of a digraph $\Gamma$ and $Q$ is an orthonormal matrix whose columns span $S$.
We reference the set $W(Q^{T}LQ)$ as the Q numerical range of the graph Laplacian $L$.
In this section, we show that the Q numerical range can be used to characterize digraphs and analyze the values of $\alpha(\Gamma)$ and $\beta(\Gamma)$. 

%%%%%%%%%%%%%%%%%%%%%%
%			Theorem 2.1			%
%%%%%%%%%%%%%%%%%%%%%%
\begin{theorem}
Let $\Gamma$ be a digraph with binary weights and let $L$ be the graph Laplacian of $\Gamma$.
Then, $\Gamma$ is a directed cycle on $n$ vertices if and only if
\begin{equation}\label{eq:cycle-eig}
\sigma(L)=\left\{1-e^{\iu 2\pi j/n}\colon~j=0,1,\ldots,n-1\right\}.
\end{equation}
\end{theorem}
\begin{proof}
Suppose that $\Gamma$ is a directed cycle on $n$ vertices. 
Then, the graph Laplacian $L=I-A$ is a circulant matrix with first column vector equal to
\[
[1,0,\cdots,0,-1]^{T}.
\]
As a circulant matrix with first column vector defined above, the eigenvalues of $L$ are known to satisfy~\eqref{eq:cycle-eig}. 

Conversely, suppose that the eigenvalues of the graph Laplacian $L$ satisfy~\eqref{eq:cycle-eig}.
Suppose that $0$ is the out-degree of some vertex of $\Gamma$.
Then, $L$ can be written in the form
\[
L=\begin{bmatrix} L_{1} & L_{12} & \cdots & L_{1r} \\ & L_{2} & \cdots & L_{2r} \\ & & \ddots & \vdots \\ & & & 0 \end{bmatrix},
\]
where $L_{1}\oplus L_{2}\oplus\cdots\oplus L_{r-1}$ is a non-singular matrix with eigenvalues 
\[
\left\{1-e^{\iu 2\pi j/n}\colon~j=1,\ldots,n-1\right\}.
\]
For each $k=1,\ldots,r-1$, by~\cite[Lemma 6.4.1]{Berman1994} and~\cite[Theorem 6.2.3 (M35)]{Berman1994}, $L_{k}$ is an irreducible non-singular M-matrix. 
Hence, by~\cite[Theorem 6.2.3 (N38)]{Berman1994}, $L_{k}^{-1}$ exists, is non-negative, and is irreducible.
Now, let $L_{k}$ denote the block with eigenvalue $1-e^{\iu 2\pi/n}$.
Then, by the Perron-Frobenius theorem~\cite[Theorem 2.1.4]{Berman1994}, it follows that $\rho(L_{k}^{-1})$ is a simple eigenvalue.
Since, $1-e^{\iu 2\pi/n}$ is the smallest eigenvalue of $L_{k}$ in magnitude, it follows that $\rho(L_{k}^{-1})=\abs{1/(1-e^{\iu 2\pi/n})}$ is an eigenvalue of $L_{k}^{-1}$. 
However, this is a contradiction and it follows that $0$ is not the out-degree of any vertex of $\Gamma$.

Since $\tr{L}=n$ and the weights of $\Gamma$ are binary, it follows that $d^{+}(i)=1$ for all $i=1,\ldots,n$.
Therefore, $L=I-A$, where the adjacency matrix $A$ has eigenvalues
\[
\left\{e^{\iu 2\pi j/n}\colon~j=0,1,\ldots,n-1\right\}.
\]
Note that $\tr{A^{k}}$ is equal to the number of closed directed walks of length $k$.
Furthermore, $\tr{A^{k}}=0$ if $k<n$ and $A^{n}=I$.
It follows that $\tr{A^{n}}=n$, so there are $n$ directed walks of length $n$.
In fact, there is a walk starting and ending at each vertex.
Since each walk must be in the same direction, it follows that $\Gamma$ is a cycle on $n$ vertices. 
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%
%			Lemma 2.2			%
%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}\label{lem:normal}
Let $L$ be the graph Laplacian of any digraph.
Then, $L$ is normal if and only if $Q^{T}LQ$ is normal.
\end{lemma}
\begin{proof}
Suppose $L$ is a normal matrix.
Then, $L$ has an orthonormal eigenvector basis for $\mathbb{C}^{n}$, which we denote by $\{v_{1},\ldots,v_{n}\}$, where $v_{n}=e/\sqrt{n}$.
Since the columns of $Q$ are orthogonal to $e$, it follows that for each $j=1,\ldots,n-1$ there is a unique $x_{j}\in\mathbb{C}^{n-1}$ such that $Qx_{j}=v_{j}$.
Therefore, $\{x_{1},\ldots,x_{n-1}\}$ forms an orthonormal basis for $\mathbb{C}^{n-1}$, where each $x_{j}$ is an eigenvector of $Q^{T}LQ$.
Hence, $Q^{T}LQ$ is normal. 

Now, suppose that $Q^{T}LQ$ is normal.
Then, $Q^{T}LQ$ has an orthonormal eigenvector basis for $\mathbb{C}^{n-1}$, which we denote by $\{x_{1},\ldots,x_{n-1}\}$.
Define $v_{j}=Qx_{j}$ for $j=1,\ldots,n-1$, which is a set of $(n-1)$ orthonormal vectors in $\mathbb{C}^{n}$.
Furthermore, each $v_{j}$ is orthogonal to $e$.
Let $v_{n}=e/\sqrt{n}$.
Then, $\{v_{1},\ldots,v_{n}\}$ forms an orthonormal basis for $\mathbb{C}^{n}$, where each $v_{j}$ is an eigenvector of $L$.
Hence, $L$ is normal. 
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%
%			Theorem 2.3			%
%%%%%%%%%%%%%%%%%%%%%%
\begin{theorem}\label{thm:cycle-qnr}
Let $\Gamma$ be a directed cycle with binary weights and let $L$ be the graph Laplacian of $\Gamma$.
Then, $\Gamma$ is a directed cycle if and only if $W(Q^{T}LQ)$ is polygon with vertices
\[
\left\{1-e^{\iu 2\pi j/n}\colon~j=1,\ldots,n-1\right\}.
\]
\end{theorem}
\begin{proof}
Suppose that $\Gamma$ is a directed cycle.
Then, the graph Laplacian $L=I-A$ is a circulant matrix with first column vector equal to
\[
[1,0,\ldots,0,-1]^{T}.
\]
As a circulant matrix, $L$ has eigenvectors that form an orthonormal basis for $\mathbb{C}^{n}$ and is therefore normal.
Furthermore, since the first column vector of $L$ is equal to the vector above, it follows that the eigenvalues of $L$ satisfy~\eqref{eq:cycle-eig}. 
Therefore, by Lemma~\ref{lem:normal}, $Q^{T}LQ$ is a normal matrix with eigenvalues that satisfy
\[
\sigma(Q^{T}LQ)=\{1-e^{\iu 2\pi j/n}\colon~j=1,\ldots,n-1\}.
\]
Hence, $W(Q^{T}LQ)$ is equal to the convex hull of $\sigma(Q^{T}LQ)$. 

Conversely, suppose that $W(Q^{T}LQ)$ is a polygon with vertices
\[
\{1-e^{\iu 2\pi j/n}\colon~j=1,\ldots,n-1\}.
\]
Then, these $(n-1)$ vertices are the eigenvalues of $Q^{T}LQ$.
Therefore, the eigenvalues of $L$ satisfy~\eqref{eq:cycle-eig}, and it follows that $\Gamma$ is a cycle on $n$ vertices. 
\end{proof}

Finally, the corollary below follows from Theorem~\ref{thm:cycle-qnr} and~\eqref{eq:alpha} and~\eqref{eq:beta}.
%%%%%%%%%%%%%%%%%%%%%%
%			Corollary 2.4			%
%%%%%%%%%%%%%%%%%%%%%%
\begin{corollary}\label{cor:cycle-alpha-beta}
If $\Gamma$ is a directed cycle on $n$ vertices, then
\[
\alpha(\Gamma) = 1-\re{e^{\iu 2\pi/n}}
\]
and
\[
\beta(\Gamma) = \begin{cases} 
				2 & \text{if $n$ is even}, \\
				1 - \re{e^{\iu \pi (n-1)/n}} & \text{if $n$ is odd}.
				\end{cases}
\]
\end{corollary}

%%%%%%%%%%%%%%%%%%%%%%
%			Remark 2.5			%
%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}~\\
\begin{itemize}
\item	The result in Corollary~\ref{cor:cycle-alpha-beta} corresponds to the values in~\cite{Wu2005-1}, which are stated without proof.
\item I believe that we can generalize Theorem 2.1 and, therefore, Theorem 2.3 to all digraphs with non-negative weights.
	We only need to work on the ``only if'' direction of Theorem 2.1.
\item	It is worth investigating other digraphs that can be characterized by the Q numerical range of their graph Laplacian.
	For instance, both the empty digraph and completely connected digraph are characterized by the Q numerical range of their graph Laplacian.
\item	Of course, the above remark will have more importance in literature if there are digraphs who are characterized by their Q numerical range but not their spectrum, or vice-versa. 
	Thus, our first goal is to find digraphs that satisfy the above remark, or prove that a digraph being characterized by its Q numerical range is equivalent to it being characterized by its spectrum. 
	I conjecture that the imploding star (see~\cite{Wu2005-1}) is characterized by its Q numerical range, but is not characterized by its spectrum.
\item My hope was that the $Q$ numerical range would characterize the complete dominance graph.
	This would help us investigate the algebraic connectivity of complete dominance graphs and applications to the rankability of data.
	Moreover, this would appease the above remark. 
	However, the following digraph is a contradiction:
	Let $\Gamma$ be defined by the graph below.
	\begin{figure}[h]
	\centering
	\resizebox{0.20\textwidth}{!}{% Non-complete Dominance Graph
	\begin{tikzpicture}
		\node[circle, shading=ball, ball color=black!75, color=white] (1) at (0,1) {\textbf{1}};
		\node[circle, shading=ball, ball color=black!75, color=white] (2) at (1,0) {\textbf{2}};
		\node[circle, shading=ball, ball color=black!75, color=white] (3) at (-1,0) {\textbf{3}};
	
		% vertex 1
		\draw[black!75,->,thick](1) to [in=45,out=225,looseness=0](3);
		% vertex 2
		\draw[black!75,->,thick](2) to [in=315,out=135,looseness=0](1);
		% vertex 3
		\draw[black!75,->,thick](3) to [in=225,out=45,looseness=0](1);
	\end{tikzpicture}%
	}
	\end{figure}
	Let $L$ be the graph Laplacian of $\Gamma$.
	Then,
	\[
	\sigma(L)=\{2,1,0\},
	\]
	which is the same spectrum as any complete dominance graph on $3$ vertices. 
	Furthermore, the singular values of $Q^{T}LQ$ are
	\[
	\sigma_{1}=\sqrt{\frac{2}{3}\left(4+\sqrt{7}\right)}~\text{ and }~\sigma_{1}=\sqrt{\frac{2}{3}\left(4-\sqrt{7}\right)}.
	\]
	Therefore, by previous note and the elliptical range theorem, the Q numerical range of $L$ is equal to the Q numerical range of the graph Laplacian associated with any complete dominance graph on $3$ vertices.
\item	Despite the remark above, we still have the worthwhile open problem of showing that if $\Gamma$ is a complete dominance graph, then the Q numerical range is an ellipse with foci equal to () and minor axis equal to (). 
\end{itemize}
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                 	   			Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{Bibliography}
\bibliographystyle{siam}
\bibliography{Bibliography}

\end{document}